---
title: "Coding_challenge7"
author: "Kiranmayee"
date: "2026-03-19"
output: 
 html_document:
      keep_md: yes
      
---

###1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
#STAND$Treatment <- as.factor(STAND$Treatment) # example shown here. 



``` {r}
STAND <- read.csv("./Coding Challenge 7/PlantEmergence.csv")

#variables into factors
STAND$Treatment <- as.factor(STAND$Treatment)
STAND$DaysAfterPlanting <- as.factor(STAND$DaysAfterPlanting)
STAND$Rep <- as.factor(STAND$Rep)

str(STAND)


```


###2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 



```{r}
lm.full <- lm(Emergence ~ Treatment + DaysAfterPlanting + Treatment:DaysAfterPlanting, 
              data = STAND)

summary(lm.full)
anova(lm.full)

```




###3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

```{r}
lm.simple <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = STAND)

summary(lm.simple)
anova(lm.simple)

```


#### A. The interaction term between Treatment and DaysAfterPlanting was not significant (P = 1.00).The simplified linear model showed that both Treatment and DaysAfterPlanting remained significant predictors of emergence.

###The intercept is 182.163. This represents the predicted emergence for the reference group, which is Treatment 1.The coefficient for Treatment 2 is -134.531. This means that Treatment 2 had 134.5 fewer emerged plants than Treatment 1 after considering DaysAfterPlanting. So,Treatment 2 had much lower emergence than Treatment 1.


###4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
install.packages("multcompView", repos = "https://cloud.r-project.org")
library(multcompView)

lsmeans.trt <- emmeans(lm.simple, ~ Treatment)
results_lsmeans <- cld(lsmeans.trt, alpha = 0.05, details = TRUE)

lsmeans.trt
results_lsmeans

```

####A.Treatment 2 had significantly lower emergence than all other treatments, with a least square mean of 55.6 emerged plants. Treatments 3, 5, and 6 had the highest emergence values, but they were not significantly different from other treatments because they share grouping letters. 

###5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) 
  
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

```

```{r}
plot_cldbars_onefactor(lm.simple, "Treatment")

```


####A.Treatments that share at least one letter are not significantly different from each other, while treatments with completely different letters are significantly different at α = 0.05. For example, Treatment 2 has the letter “c” and is significantly different from all other treatments, while Treatments 3, 5, and 6 all have “a” and are not significantly different from one another.


###6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

https://github.com/kzb0148/PLPA_2026/tree/main/Coding%20Challenge%207

